---
title: "Visualizing 2016 Primary Election Result"
author: Wenqi Cheng, Andrew Cho, Kshipra Hemal, Liaowang Zou
output: html_document
runtime: shiny
---
### Project Overview
Our Shiny app displays the most recent election primaries results in a US State map. User can select a presidential candidate and view which states they have won and a chloropleth map that color-projects each candidate's popularity in each state. There is also an option to view the current delegate count for each of the political parties. 

### Part I - Scrape / Parse 
We read this url, "http://www.politico.com/2016-election/results/map/president".  We used "Selector Gadget" to generate a minimal CSS selector to get five candidates' delegate and percentage data. Since we got a long vector, containing all data. We splited it by state and created a data frame df to store all data, including state names, delegate, percentage, winners' names.
```{r}
suppressMessages(library(knitr))
load("df0.Rdata")
kable(head(df0))
```

### Part II- Make a Shiny Chloropleth Map
The main output for this Shiny App is a chloropleth map of the United State. Users can first filter by parties, then select a candidate of interest. The map would display the percentage of support for that candidate in 2016 primary election by state. The darker the color, the higher the percentage of support. We also show the states in which the candidate won on top of the map. Finally, we include bar plots showing how many more delegates the candidates need in order to be nominated are also displayed. Because we choose the scrape the data from Politico's website, our App will automatically update its results whenever Politico updates.

We first create the dataframes for the barplot of delegates. Because they are not reactive and does not depend on the user's input, we can create them outside of the Shiny App. To obtain the total delegate number for each candidate, we sum the number of delegates by state for each candidate. Note here we used the original dataframe `df0` instead of `df`, which is altered in order to fit the order of the "states" in map function. 

To creating a chloropleth map , we use a helper function `percent_map` that utilize the map function in "maps" package. Our code is adapted from the Shiny tutorial [Link](http://shiny.rstudio.com/tutorial/lesson5/). 

Finally, we piece every component together and create the Shiny App. Our work product is displayed below:

```{r}
## Load any packages 
suppressMessages(library(shiny)) 
suppressMessages(library(mapproj))
suppressMessages(library(magrittr))
suppressMessages(library(dplyr))
suppressMessages(library(stringr)) 
suppressMessages(library(maps))

# load scraped data
load("df0.Rdata")
load("df.Rdata")

## dataframes for running barplot (democrats)
dem = c(sum(df0$Dele_Clinton,na.rm = TRUE),sum(df0$Dele_Sanders,na.rm = TRUE))
names(dem) = c("Clinton", "Sanders")

## dataframes for running barplot (republican)
rep = c(sum(df0$Dele_Trump,na.rm = TRUE),sum(df0$Dele_Cruz,na.rm = TRUE),sum(df0$Dele_Kasich,na.rm = TRUE))
names(rep) = c("Trump", "Cruz", "Kasich")

## helper function in outputting the map
percent_map = function(var, color, legend.title, min = 0, max = 100) {
  # generate vector of fill colors for map
  shades = colorRampPalette(c("white", color))(100)
  # constrain gradient to percents that occur between min and max
  var = pmax(var, min)
  var = pmin(var, max)
  percents = as.integer(cut(var, 100, 
                            include.lowest = TRUE, ordered = TRUE))
  fills = shades[percents]
  # plot choropleth map
  par(mar = c(0,0,0,0))
  map("state", fill = TRUE, col = fills, 
      resolution = 0, lty = 0, projection = "polyconic", 
      myborder = 0, mar = c(0,0,0,0))
  # overlay state borders
  map("state", col = "gray", fill = FALSE, add = TRUE,
      lty = 1, lwd = 1, projection = "polyconic", 
      myborder = 0, mar = c(0,0,0,0))
  # add a legend
  inc = (max - min) / 4
  legend.text = c(paste0(min, " % or less"),
                  paste0(min + inc, " %"),
                  paste0(min + 2 * inc, " %"),
                  paste0(min + 3 * inc, " %"),
                  paste0(max, " % or more"))
  legend("bottomleft", 
         legend = legend.text, 
         fill = shades[c(1, 25, 50, 75, 100)], 
         title = legend.title, bty = "n")
}

## App

shinyApp(
  ui = fluidPage(
    tags$head(
      tags$style(".leaflet-container { background-color: white !important; }"
      )
    ),
    
    # user first chooses Party, then the party affiliated candidate
    sidebarPanel(
      selectInput(inputId = "Party", label = "Select a political party:", 
                  choices = c("Democratic", "Republican"),
                  selected = "Democratic"),
      br(),
      uiOutput("candidate"), # output based on Party selection by user
      # toggle on/off plot of running delegate counts
      checkboxInput("yes_demdel", "Show Democratic Delegate Count", value = TRUE), 
      checkboxInput("yes_repdel", "Show Republican Delegate Count", value = TRUE)
    ),
    
    mainPanel(
      textOutput("cand_won"),
      tags$head(tags$style("#cand_won{font-size: 16px; 
                                      font-weight: bold;
                                      }"
                           )
                ),
      br(),
      textOutput("states_won"),
      plotOutput("map"),
      conditionalPanel(
        condition = "input.yes_demdel == true",
        plotOutput("demdel")
      ),
      conditionalPanel(
        condition = "input.yes_repdel == true",
        plotOutput("repdel")
      )
    )
  ),  
  
  server = function(input, output, session){
    # returns names of candidates based on user chosen party affiliation
    candidates = reactive({
      if(input$Party == "Republican"){
        return(c("Trump", "Kasich", "Cruz"))
      } else {
        return(c("Clinton", "Sanders"))
      }
    })
    
    # takes user chosen party and displays corresponding candidates
    output$candidate = renderUI({
      selectInput("cand", "Candidate", 
                  choices = candidates())
    })
    
    # outputs states that a candidate has won
    states = reactive({
      if(input$Party=="Democratic"){
        return(df$state2[str_detect(df$D_win, input$cand)])
      } else{
        return(df$state2[str_detect(df$R_win, input$cand)])
      }
    })
    
    output$cand_won = renderText({
      paste("States that ", input$cand, "has won: ")
    })
    
    output$states_won = renderText({
      st = unlist(unique(states())) %>%
        sapply(function(x) paste0(toupper(substr(x,1,1)), substr(x,2,nchar(x)))) 
      paste(st,collapse = ", ")
    })
    
    
    # map of the voter results for a candidate. Darker color corresponds to higher % votes
    output$map = renderPlot({
      if (input$Party == "Democratic"){
        percent_map(df[[tolower(input$cand)]], color = "royalblue", input$cand)
      } else { # if Republican
        percent_map(df[[tolower(input$cand)]], color = "red", input$cand)
      }  
    })
    
    # displays Republican candidate delegate counts
    output$repdel = renderPlot({
      par(mar = c(6,6,2,1))
      graphrep = barplot(rep,main = "Republican Delegates", 
              space= 0.5,
              col=c("orange", "maroon", "gold"), xpd=T, cex.names = 0.95, xaxt = "n", las = 1,
              border = NA, horiz = TRUE, xlim = c(0, 1500), ylim = c(0,1), width = rep(0.2, 3)) 
      text(x = 80, y = graphrep, label = rep, cex= 1.5)
      abline(v=1237, col = "red", lty = 2)
      legend("bottomleft", col = "red", lwd = 2, lty = 2,
              legend = "*Delegates needed to win party nomination (1237)", 
              bty = "n")
    })
    
    # displays Democratic candidate delegate counts 
    output$demdel = renderPlot({
      par(mar = c(6,6,2,1))
      graphdem = barplot(dem, main = "Democratic Delegates", 
              col=c("sky blue", "pale green"), cex.names = 0.95, xaxt = "n", las = 1, space = 0.5,
              border = NA, horiz = TRUE,  xlim = c(0,2500), ylim = c(0,1), width = rep(0.25, 2))
      text(x = 150, y = graphdem , label = dem, cex = 1.5)
      abline(v=2383, col = "red", lty = 2)
      legend("bottomleft", col = "red", lwd = 2, lty = 2,
              legend = "*Delegates needed to win party nomination (2383)", bty = "n")
    })
  },
  options = list(height = 500)
)
```

